Forecasting the 2024 U.S. Presidential Election: Kamala Harris' Projected Victory

### 2024-11-04 paper/paper.qmd
committing changes to appendix for paper.qmd and testing scripts

---
title: "Forecasting the 2024 U.S. Presidential Election: Kamala Harris' Projected Victory"
subtitle: A Bayesian Analysis of Polling Data from Key Battleground States"
author: 
  - Jimin Lee
  - Sarah Ding
  - Xiyan Chen
thanks: "Code and data are available at: [https://github.com/jamiejiminlee/2024_US_Elections.git](https://github.com/jamiejiminlee/2024_US_Elections.git)."
date: today
date-format: long
abstract: "This paper presents a predictive model for the 2024 United States Presidential Election, focusing on essential battleground states. Utilizing a Bayesian framework and state-level polling data, we estimate the probabilities of victory for candidates Donald Trump and Kamala Harris. Our findings indicate that Harris is projected to secure a win in six out of seven key battleground states, such as Michigan, Nevada, and North Carolina, with predicted support exceeding 50%. This analysis not only forecasts the likely winner of the U.S. election but also equips society and the economy to better undertstand the current U.S. political climate."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

# Introduction {#sec-intro}
The 2024 United States Presidential Election is set to be a crucial moment in American political history, with battleground states serving as pivotal determinants of the electoral outcome. As candidates Donald Trump and Kamala Harris intensify their campaigns, understanding voter preferences in these key regions is essential for accurate election forecasting. This paper utilizes state-level polling data from FiveThirtyEight [@FiveThirtyEight] and the statistical programming language R [@citeR] to develop a predictive model estimating each candidate's likelihood of victory in critical battleground states, including Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin.

@@

These predictions are critical not only for forecasting the likely outcome of the election but also for enhancing societal and economic understanding of the current political climate in the U.S. The implications of our findings extend beyond mere predictions, offering a framework for analyzing potential shifts in voter sentiment as Election Day approaches. The remainder of this paper is organized as follows: Section @sec-data details the data and variables utilized in the analysis; Section @sec-model outlines the model setup and estimation strategies employed; Section @sec-results presents the results, including visualizations of the predicted election outcomes; and Section @sec-discussion discusses the implications and limitations of our findings.

All programming for this paper were conducted using the statistical programming language `R` [@citeR] and the following packages: `tidyverse` package of @Tidyverse , `janitor` package of @Janitor for data cleaning, `kableExtra` package of @kableExtra for dynamic report generation and table formatting, `modelsummary` package of @Modelsummary for summarizing models, `usmap` package of @usmap for mapping U.S. states, `arrow` package of @arrow for data integration with Apache Arrow, `lubridate` package of @Lubridate for date handling, `rstanarm` package of @rstanarm and `brms` package of @brms for Bayesian regression modeling, `Matrix` package of @Matrix for handling sparse and dense matrices, `splines` package of @splines for regression splines, `caret`package of @caret for machine learning support, and `broom.mixed` package of @broommixed for tidying mixed models.

# Data {#sec-data}
## Overview
The data for this paper contains the polling information of the 2024 Presidential Election cycle and was obtained from Project 538's polling project [@FiveThirtyEight] on October 19th, 2024. The data is periodically updated throughout the presidential election campaign to reflect current and most up-to-date polling results. It consists of polling information from various pollsters such as the New York Times, ActiVote, Morning Consult, etc. For every poll conducted, relevant variables such as **'state'**, **'start_date'**, **'end_date'**, **'transparency_score'**, **'candidate_name'**, and percentage of support (**'pct'**) for the appropriate candidates are included. The data compiles information from all major pollsters in the US and provides us with an understanding of the current state of the presidential election.

The present analysis focuses on Donald Trump and Kamala Harris, the two leading candidates in the current presidential election race. For all the variables mentioned (see @sec-variables), we collected observations for both of the candidates. We collect observations from the data where the **‘candidate_name’** is Trump and Harris. We are interested in the support for both candidates over time, for each state, therefore, we collect observations of **‘end_date’**, and **‘state’**, as well as the electorate support for each candidate is denoted as **‘pct’**. Furthermore, since there are many pollsters present in the data, we extracted pollsters with high-quality data that is reliable, which is denoted as having a high **‘numeric_grade’** and **‘transparency_score’**, details of data cleaning are further explained in [Appendix -@sec-data-details].

@@

- **'State'**: Indicates the state where the polling took place, critical for understanding regional support dynamics.
- **'Polling Date'**: The date when the poll was conducted, which helps in analyzing trends over time.
- **'Pollster'**: The organization conducting the poll, providing insight into the reliability and methodology of the polling data.
- **'Numeric Grade'**: A score assigned to each poll based on its methodology and historical accuracy, which assists in filtering out less reliable polls.
- **'Transparency Score'**: A measure indicating how transparent the pollster is regarding their methodology, further enhancing data credibility.
- **'Candidate Name'**: Identifies the candidate being polled, specifically focusing on Donald Trump and Kamala Harris.
- **Percentage of Support (pct)**: Represents the percentage of respondents supporting each candidate, which is the primary outcome variable of interest.

@@

### Outcome Variable
As our focus is to analyze and predict voter support for Kamala Harris and Donald Trump during the current presidential election cycle, the primary outcome variables of interest are the percentages of support for each candidate, denoted as ‘pct’.The pct is categorized by candidate and scaled to a proportion by diving by 100. This variable represents the proportion of voter support for each candidate based on polling data. Through understanding the percentage of support (pct) for each candidate, we can observe and infer the candidate that has a leading advantage in winning the presidential election race.

@@

```
We filtered for the earliest end date as July 2024 since Harris officially announced her presidential campaign on July 21, 2024. Therefore, any polls that has an end date before July would not contain polling results for Harris. The data from [@FiveThirtyEight] is last updated on October 15, 2024, thus the latest end date being October 21, 2024. Furthermore, @fig-end-date-distribution also reveals that polling activity is not evenly distributed over time. There are clear peaks where the number of polls conducted is higher, notably in early September, late September, as well as early October. This concentration of polls likely corresponds to period of heightened political interest or significant campaign events that prompt increased polling.

@@

#### Numeric Grade
Numeric Grade (**'numeric_grade'**) represents the quality of each poll, as rated by [@FiveThirtyEight], based on factors such as methodology, transparency, and historical accuracy. In the context of our paper, numeric grade allows us to weigh the reliability of different polls. By including numeric grade as a variable in our model, we aim to account for variations in poll quality, ensuring that polls with higher ratings (indicating higher reliability) contribute more significantly to our predictions of voter support. 
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-numgrade-distribution
#| fig-cap: "Density plot illustrating the distribution of numeric grades assigned to pollsters within the polling data. Highlights the frequency of numeric grades, indicating key concentrations that suggest variability in pollster quality. "

@@

#### Transparency Score
Transparency Score (**'transparency_score'**) represents the degree of methodological transparency provided by each pollster, rated on a scale from 0 to 10 by [@FiveThirtyEight]. This score reflects the extent to which pollsters disclose details about their data collection and analysis methods, such as sample size, weighting, and margin of error. In the context of our paper, transparency score is an essential variable for evaluating the reliability of polling data, as higher transparency scores generally indicate a greater level of methodological rigor and openness. @fig-transparency-score-distribution displays the distribution of transparency scores across the polling data for key battleground states. The plot shows that transparency scores cluster heavily around ratings of  7 and 9, suggesting that a majority of polls in our dataset come from sources that maintain a relatively high level of transparency. There are smaller peaks at lower scores (5 and 10), indicating that while some pollsters offer limited transparency, a few polls also achieve perfect transparency scores.
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-transparency-score-distribution
#| fig-cap: "Histogram displaying the distribution of transparency scores among pollsters in the polling data. Indicates the frequency of various transparency scores, highlighting a concentration of scores around 7 and 9, which suggests a predominance of high transparency among the assessed pollsters. "

@@

# Horizontal bar graph for the distribution of pollsters
ggplot(clean_poll_data, aes(x = factor(pollster))) +  # Convert to factor to ensure proper categorization
  geom_bar(fill = "purple") +
  labs(
       x = "Count of Polls",
       y = "Pollster") +  
  theme_minimal() +
  coord_flip() +  
  theme(axis.text.y = element_text(angle = 0, hjust = 1))  
```
@fig-pollster-distribution displays the count of polls conducted by each pollster, illustrating the distribution of polling efforts across multiple organizations. Notably, Emerson is the most frequent pollster, conducting the highest number of polls, followed by Quinnipiac and Siena/NYT. Other pollsters, such as YouGov, CNN/SSRS, and Marist, also contribute a significant number of polls, whereas some organizations like The Washington Post and SurveyUSA have comparatively fewer entries in the dataset.This distribution is crucial for evaluating the consistency and breadth of the polling data. Pollsters with a higher count of polls, like Emerson, contribute substantially to the dataset and can have a greater influence on the model's predictions, particularly if their methodology or sample selection differs from others. On the other hand, pollsters with fewer entries may provide valuable but limited data points, potentially adding unique perspectives but with less influence on overall trends.

@@

## Model Overview
The Bayesian Beta regression models in this paper estimate the probability of victory for Donald Trump and Kamala Harris in the 2024 U.S. Presidential Election by analyzing polling data from seven battleground states. Separate models are constructed for each candidate to estimate their support, considering polling trends over time and state-level differences in voter sentiment. This approach captures how support fluctuates leading up to Election Day and predicts potential electoral outcomes. To analyze changes in voter support over time, the models use natural splines with the **'end_date_num'** (days from the earliest poll date) as a predictor, allowing for the identification of non-linear trends influenced by campaign events and news cycles. Additionally, state-level effects are incorporated as random factors, enabling the models to account for unique political landscapes across the battleground states. This regional adjustment helps highlight variations in candidate support and enhances the accuracy of predictions. Further detailed model specifications and diagnostics are provided in [Appendix -@sec-model-details].

@@

## Model Justification
The Bayesian Beta Regression model with a logit link function is the most suitable model for predicting voter support for Kamala Harris and Donald Trump in the 2024 U.S. Presidential Election. Our primary outcome variable is the percentage of support for each candidate in seven key battleground states, which is naturally bounded between 0% and 100%. The Beta distribution is ideal for modeling this proportion data, as it allows us to avoid unrealistic predictions outside the (0, 1) range. To capture non-linear trends in voter support as Election Day approaches, we employe natural splines with **end_date_num**, which represents the number of days since the earliest poll date. Additionally, incorporating state-specific random effects acknowledges the variations in voter preferences across different regions, allowing for more accurate modeling of baseline support levels. The Bayesian framework enhances our model by enabling the incorporation of prior information, leading to more stable estimates, especially in the context of sparse polling data. Furthermore, the model's structure allows for clear interpretation in terms of victory probabilities, aligning well with our goals of forecasting electoral outcomes. Further details are discussed in [Appendix -@sec-model-details].

@@

## State-Level Polling Averages for Trump and Harris
@tbl-stateaverage displays overview of the percentage of support for Trump and Harris by each battleground state. It provides us with a high level understanding of the current voter dynamic. The pct results are aggregated averages of all pct data points for Trump and Harris in each state. By comparing the percentage of support for T.rump and Harris, we can determine the candidate that has a leading advantage in winning each state. For example,Trump has 49.2% support in Arizona while Harris has 46.8% support, which means Trump has won the state of Arizona.
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-stateaverage
#| tbl-cap: "Percentage of support for Donald Trump and Kamala Harris across key battleground states. Trump maintains a slight lead in states such as Arizona and Georgia, while Harris shows stronger support in 5 of 7 key battleground states."

@@

 # Render the table
 knitr::kable(average_table)
```
This distribution is crucial for evaluating the consistency and breadth of the polling data. Pollsters with a higher count of polls, like Emerson, contribute substantially to the dataset and can have a greater influence on the model's predictions, particularly if their methodology or sample selection differs from others. On the other hand, pollsters with fewer entries may provide valuable but limited data points, potentially adding unique perspectives but with less influence on overall trends.

## Polling Averages for Trump and Harris Over Time in Battleground States
@fig-state-polling-average provides a visual comparison of the current polling support for Donald Trump and Kamala Harris across key battleground states, tracked over time from August 2024 to October 2024. Each line plot represents the average polling percentages for both candidates in a specific state, with red lines denoting Trump’s support and blue lines representing Harris’s support. Across several states, there are noticeably increased fluctuations in polling averages for both candidates closer to the election particularly from late September to October. This pattern may indicate heightened voter interest and engagement as the election date approaches which is November 4th, 2024, it could also reflect changes in public opinion due to campaign events, political debates, or other external factors.
```{r}
#| echo: false
#| eval: true
#| label: fig-state-polling-average
#| fig-cap: "Polling averages for Donald Trump (red) and Kamala Harris (blue) in battleground states from August to October 2024. Captures the varying fluctuations in average support percentages across Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin, indicating the competitive dynamics of the election."
#| warning: false

@@

# Print results in table
kable(state_predictions, 
      col.names = c("State", 
                    "Predicted Trump Win (%)", 
                    "Predicted Harris Win (%)", 
                    "Predicted Winner")) %>%
  kable_styling("striped", full_width = F)

```

## US Map Showing Predicted Winner by State on Election Day
To visualize the colour distribution of the predicted results, we emloy the US map from [@usmap] to display the results from @tbl-state-probabilities.

@@

## Uncertainty Beyond Data and Modeling
Beyond data and modeling, external factors such as recency bias, social desirability bias, and non-response bias can influence the election outcome. Recency bias significantly shapes public perception, with many expecting Trump to win due to his 2016 victory and narrow 2020 loss, despite lagging in polls. This bias highlights a tendency to focus on recent events while overlooking broader trends. For example, in 2012, Barack Obama outperformed polls, a fact often overshadowed by recent election narratives[@discussion-article3]. This can skew how people respond to polls and how pollsters interpret results, introducing potential polling errors that our model cannot fully account for.

Social desirability bias and non-response bias add further uncertainty, especially regarding Trump’s support base. Social stigma may deter some voters from expressing conservative support openly, leading to under-reported Trump support in polls. Additionally, Trump’s supporters often have lower social trust, making them less likely to participate in polls altogether[@discussion-article3]. If certain groups systematically avoid participation, polling data may under-represent these voters, resulting in an underestimation of Trump’s support in our model.

Harris faces potential challenges such as the Bradley effect and gender-related bias. As a Black female candidate, Harris may encounter undecided voters leaning against her due to underlying biases, similar to what Hillary Clinton faced in 2016 [@discussion-article3].  If undecided voters ultimately vote against Harris, our model may overestimate her support based on polling data alone, affecting prediction accuracy.

@@

## Model Diagnostics {#sec-model-diagnostics}
To assess the validity and reliability of our Bayesian models for predicting voter support for Kamala Harris and Donald Trump, we conducted posterior predictive checks for both models. These checks are essential for evaluating the models' ability to replicate observed polling data based on the estimated parameters. By comparing the distributions of observed support percentages with the predicted values, we can identify discrepancies and assess the overall fit of our models. Posterior predictive checks help verify model assumptions and provide insights into how well the models capture the underlying voter dynamics. A close alignment between observed and predicted values suggests that our models effectively reflect voter support trends in the electoral context. In contrast, significant deviations may reveal areas for model improvement or highlight limitations in our approach. This diagnostic process aims to enhance the credibility of our findings and deepen our understanding of the electoral landscape as the 2024 Presidential Election approaches

@@

### Strengths of the NYT/Siena College questionnaires
First, they have a comprehensive question design. NYT/Sienna College polls ask a wide range of questions that provides a detailed view of the respondent’s preferences and opinions. It helps produce multi-dimensional data that allows for in-depth analysis and segmentation of voter groups. 
Second, the questionnaires are tailored to likely voters, as identified through the L-2 voter file. This means that the voters are more likely to be up-to-date with relevant policies proposed by different parties, and have formed their own opinions about them, this improves the predictive reliability of the poll. 
Third, the questionnaires contain detailed demographic questions about the respondents which allow NYT/Siena College to stratify responses by voter subgroups, identifying varied voter bases and diverse issues.

### Weaknesses of the NYT/Siena College questionnaires
First, the length and complexity of the questionnaires might cause the respondents to be less likely to respond or finish the polls. Respondents may rush through questions or drop out before completion. Respondents may also be prone to moderacy response bias where they have the tendency to choose middle options regardless of question content, as well as response order bias where respondents choose answers based on their order. 
Second, questionnaires have a dependence on self-reported likelihood to vote. Self-reported likelihood can be unreliable, especially for individuals whose intentions to vote may change close to the actual election date. This can reduce the accuracy of predictions based on likely voter models. 
Third, there are limited open-ended questions. Many voters may have mixed opinions on specific issues or policies that cannot be explained in a single-choice answer. However, answers to open-ended questions are often difficult to convert to actionable data, so questionnaires tend to rely on closed-ended questions for scalability. This might restrict the ability to capture complex reasonings behind their choices, which could add extra strain on the time and resources used to interpret the answers.

# Idealized Methodology
Building on our discussion of the New York Times/Siena College Poll, we now present an idealized methodology that aims to enhance the accuracy and reliability of our predictions. This approach incorporates best practices and innovative techniques to ensure a comprehensive prediction of voter behavior in the upcoming 2024 U.S. presidential election. To demonstrate our idealized methodology, we generated a survey on the 2024 US Presidential election, which can be accessed via the the URL provided in [@US-election-survey] under [References -@sec-references].

>>>>

---
title: "Forecasting the 2024 U.S. Presidential Election: Kamala Harris' Projected Victory"
subtitle: "A Bayesian Analysis of Polling Data from Key Battleground States - Estimating Voter Support Dynamics for Donald Trump and Kamala Harris in the 2024 U.S. Presidential Election"
author: 
  - Jimin Lee
  - Sarah Ding
  - Xiyan Chen
thanks: "Code and data are available at: [https://github.com/jamiejiminlee/2024_US_Elections.git](https://github.com/jamiejiminlee/2024_US_Elections.git)."
date: today
date-format: long
abstract: "This paper presents a predictive model for the 2024 United States Presidential Election, focusing on essential battleground states. Utilizing a Bayesian framework and state-level polling data, we estimate the probabilities of victory for candidates Donald Trump and Kamala Harris. Our findings indicate that Harris is projected to secure a win in six out of seven key battleground states, with predicted support exceeding 50%. This analysis not only forecasts the likely winner of the U.S. election but also equips society and the economy to better undertstand the current U.S. political climate."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

# Introduction {#sec-intro}
The 2024 United States Presidential Election is set to be a crucial moment in American political history, with battleground states serving as pivotal determinants of the electoral outcome. As candidates Donald Trump and Kamala Harris intensify their campaigns, understanding voter preferences in these key regions is essential for accurate election forecasting. This paper utilizes state-level polling data from Project 538 [@FiveThirtyEight] and the statistical programming language R [@citeR] to develop a predictive model estimating each candidate's likelihood of victory in critical battleground states, including Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin.

@@

These predictions are critical not only for forecasting the likely outcome of the election but also for enhancing societal and economic understanding of the current political climate in the U.S. The implications of our findings extend beyond mere predictions, offering a framework for analyzing potential shifts in voter sentiment as Election Day approaches. The remainder of this paper is organized as follows: @sec-data details the data and variables utilized in the analysis; @sec-model outlines the model setup and estimation strategies employed; @sec-results presents the results, including visualizations of the predicted election outcomes; and @sec-discussion discusses the implications and limitations of our findings.

Data cleaning, analysis, modeling, and visualization for this paper were conducted using the statistical programming language `R` [@citeR] and the following packages: `tidyverse` package of @Tidyverse , `janitor` package of @Janitor for data cleaning, `kableExtra` package of @kableExtra for dynamic report generation and table formatting, `modelsummary` package of @Modelsummary for summarizing models, `usmap` package of @usmap for mapping U.S. states, `arrow` package of @arrow for data integration with Apache Arrow, `lubridate` package of @Lubridate for date handling, `rstanarm` package of @rstanarm and `brms` package of @brms for Bayesian regression modeling, `Matrix` package of @Matrix for handling sparse and dense matrices, `splines` package of @splines for regression splines, `caret`package of @caret for machine learning support, and `broom.mixed` package of @broommixed for tidying mixed models.

# Data {#sec-data}
## Overview
The data for this paper contains the polling information of the 2024 Presidential Election cycle and was obtained from Project 538 [@FiveThirtyEight] on October 19th, 2024. The data is periodically updated throughout the presidential election campaign to reflect current and most up-to-date polling results. It consists of polling information from various pollsters such as the New York Times, ActiVote, Morning Consult, etc. For every poll conducted, relevant variables such as **'state'**, **'start_date'**, **'end_date'**, **'transparency_score'**, **'candidate_name'**, and percentage of support (**'pct'**) for the appropriate candidates are included. The data compiles information from all major pollsters in the US and provides us with an understanding of the current state of the presidential election.

The present analysis focuses on Donald Trump and Kamala Harris, the two leading candidates in the current presidential election race. For all the variables mentioned (see @sec-variables), we collected observations for both of the candidates. We collect observations from the data where the **‘candidate_name’** is Trump and Harris. We are interested in the support for both candidates over time, for each state, therefore, we collect observations of **‘end_date’** and **‘state’**, as well as the electorate support for each candidate, which is denoted as **‘pct’**. Furthermore, since there are many pollsters present in the data, we extracted pollsters with high-quality data that is reliable, which is denoted as having a high **‘numeric_grade’** and **‘transparency_score’**, details of data cleaning are further explained in [Appendix -@sec-data-details].

@@

- **'State'**: Indicates the state where the polling took place, critical for understanding regional support dynamics.
- **'End Date'**: The date when the poll ended, which helps in analyzing trends over time.
- **'Start Date'**: The date when the poll started, to filter for Kamala Harris's late entry into the presidential race 
- **'Pollster'**: The organization conducting the poll, providing insight into the reliability and methodology of the polling data.
- **'Numeric Grade'**: A score assigned to each poll based on its methodology and historical accuracy, which assists in filtering out less reliable polls.
- **'Transparency Score'**: A measure indicating how transparent the pollster is regarding their methodology, further enhancing data credibility.
- **'Candidate Name'**: Identifies the running candidate, specifically focusing on Donald Trump and Kamala Harris.
- **Percentage of Support (pct)**: Represents the percentage of respondents supporting each candidate, which is the primary outcome variable of interest.

@@

### Outcome Variable
As our focus is to analyze and predict voter support for Kamala Harris and Donald Trump during the current presidential election cycle, the primary outcome variables of interest are the percentages of support for each candidate, denoted as **‘pct’**.The pct is categorized by candidate and scaled to a proportion by diving by 100. This variable represents the proportion of voter support for each candidate based on polling data. Through understanding the percentage of support (**'pct'**) for each candidate, we can observe and infer the candidate that has a leading advantage in winning the presidential election race.

@@

```
We filtered for the earliest end date as July 2024 since Harris officially announced her presidential campaign on July 21, 2024. Therefore, any polls that has an end date before July would not contain polling results for Harris. The data from Project 538 [@FiveThirtyEight] is last updated on October 15th, 2024, thus the latest end date being October 14th, 2024. Furthermore, @fig-end-date-distribution also reveals that polling activity is not evenly distributed over time. There are clear peaks where the number of polls conducted is higher, notably in early September, late September, as well as early October. This concentration of polls likely corresponds to period of heightened political interest or significant campaign events that prompt increased polling.

@@

#### Numeric Grade
Numeric Grade (**'numeric_grade'**) represents the quality of each poll, as rated by Project 538 [@FiveThirtyEight], based on factors such as methodology, transparency, and historical accuracy. In the context of our paper, numeric grade allows us to weigh the reliability of different polls. By including numeric grade as a variable in our model, we aim to account for variations in poll quality, ensuring that polls with higher ratings (indicating higher reliability) contribute more significantly to our predictions of voter support. 
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-numgrade-distribution
#| fig-cap: "Density plot illustrating the distribution of numeric grades assigned to pollsters within the polling data. Highlights the frequency of numeric grades, indicating key concentrations that suggest variability in pollster quality. "

@@

#### Transparency Score
Transparency Score (**'transparency_score'**) represents the degree of methodological transparency provided by each pollster, rated on a scale from 0 to 10 by Project 538 [@FiveThirtyEight]. This score reflects the extent to which pollsters disclose details about their data collection and analysis methods, such as sample size, weighting, and margin of error. In the context of our paper, transparency score is an essential variable for evaluating the reliability of polling data, as higher transparency scores generally indicate a greater level of methodological rigor and openness. @fig-transparency-score-distribution displays the distribution of transparency scores across the polling data for key battleground states. The plot shows that transparency scores cluster heavily around ratings of  7 and 9, suggesting that a majority of polls in our dataset come from sources that maintain a relatively high level of transparency. There are smaller peaks at lower scores (5 and 10), indicating that while some pollsters offer limited transparency, a few polls also achieve perfect transparency scores.
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-transparency-score-distribution
#| fig-cap: "Histogram displaying the distribution of transparency scores among pollsters in the polling data. Indicates the frequency of various transparency scores, highlighting a concentration of scores around 7 and 9, which suggests a predominance of high transparency among the assessed pollsters. "

@@

# Horizontal bar graph for the distribution of pollsters
ggplot(clean_poll_data, aes(x = factor(pollster))) +  # Convert to factor to ensure proper categorization
  geom_bar(fill = "purple") +
  labs(
       x = "Count of Polls",
       y = "Pollster") +  
  theme_minimal() +
  coord_flip() +  
  theme(axis.text.y = element_text(angle = 0, hjust = 1))  
```
@fig-pollster-distribution displays the count of polls conducted by each pollster, illustrating the distribution of polling efforts across multiple organizations. Notably, Emerson is the most frequent pollster, conducting the highest number of polls, followed by Quinnipiac and Siena/NYT. Other pollsters, such as YouGov, CNN/SSRS, and Marist, also contribute a significant number of polls, whereas some organizations like The Washington Post and SurveyUSA have comparatively fewer entries in the dataset. This distribution is crucial for evaluating the consistency and breadth of the polling data. Pollsters with a higher count of polls, like Emerson, contribute substantially to the dataset and can have a greater influence on the model's predictions, particularly if their methodology or sample selection differs from others. On the other hand, pollsters with fewer entries may provide valuable but limited data points, potentially adding unique perspectives but with less influence on overall trends.

@@

## Model Overview
The Bayesian Beta regression models in this paper estimate the probability of victory for Donald Trump and Kamala Harris in the 2024 U.S. Presidential Election by analyzing polling data from seven battleground states. Separate models are constructed for each candidate to estimate their support, considering polling trends over time and state-level differences in voter sentiment. This approach captures how support fluctuates leading up to Election Day and predicts potential electoral outcomes. To analyze changes in voter support over time, the models use natural splines with the **'end_date_num'** (days from the earliest **'end_date'** to the latest **'end_date'**) as a predictor, allowing for the identification of non-linear trends influenced by campaign events and news cycles. Additionally, state-level effects are incorporated as random factors, enabling the models to account for unique political landscapes across the battleground states. This regional adjustment helps highlight variations in candidate support and enhances the accuracy of predictions. Further detailed model specifications and diagnostics are provided in [Appendix -@sec-model-details].

@@

## Model Justification
The Bayesian Beta Regression model with a logit link function is the most suitable model for predicting voter support for Kamala Harris and Donald Trump in the 2024 U.S. Presidential Election. Our primary outcome variable is the percentage of support for each candidate in seven key battleground states, which is naturally bounded between 0% and 100%. The Beta distribution is ideal for modeling this proportion data, as it allows us to avoid unrealistic predictions outside the (0, 1) range. To capture non-linear trends in voter support as Election Day approaches, we employe natural splines with **'end_date_num'**, which represents the number of days from earliest **'end_date'** to latest **'end_date'**. Additionally, incorporating state-specific random effects acknowledges the variations in voter preferences across different regions, allowing for more accurate modeling of baseline support levels. The Bayesian framework enhances our model by enabling the incorporation of prior information, leading to more stable estimates, especially in the context of sparse polling data. Furthermore, the model's structure allows for clear interpretation in terms of victory probabilities, aligning well with our goals of forecasting electoral outcomes. Further details are discussed in [Appendix -@sec-model-details].

@@

## State-Level Polling Averages for Trump and Harris
@tbl-stateaverage displays overview of the percentage of support for Trump and Harris by each battleground state. It provides us with a high level understanding of the current voter dynamic. The percentages of support for Trump and Harris are aggregated averages of all **'pct'** data points for Trump and Harris in each state. By comparing the percentage of support for Trump and Harris, we can determine the candidate that has a leading advantage in winning each state. For example, Trump has 49.2% support in Arizona while Harris has 46.8% support, which means Trump has won the state of Arizona.
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-stateaverage
#| tbl-cap: "Percentage of support for Donald Trump and Kamala Harris across key battleground states. Trump maintains a slight lead in states such as Arizona and Georgia, while Harris shows stronger support in 5 of 7 key battleground states."

@@

 # Render the table
 knitr::kable(average_table)
```

## Polling Averages for Trump and Harris Over Time in Battleground States
@fig-state-polling-average provides a visual comparison of the current polling support for Donald Trump and Kamala Harris across key battleground states, tracked over time from August 2024 to October 2024. Each line plot represents the average polling percentages for both candidates in a specific state, with red lines denoting Trump’s support and blue lines representing Harris’s support. Across several states, there are noticeably increased fluctuations in polling averages for both candidates closer to the election particularly from late September to October. This pattern may indicate heightened voter interest and engagement as the election date approaches which is November 5th, 2024, it could also reflect changes in public opinion due to campaign events, political debates, or other external factors.
```{r}
#| echo: false
#| eval: true
#| label: fig-state-polling-average
#| fig-cap: "Polling averages for Donald Trump (red) and Kamala Harris (blue) in battleground states from August to October 2024. Captures the varying fluctuations in average support percentages across Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin, indicating the competitive dynamics of the election."
#| warning: false

@@

# Print results in table
kable(state_predictions, 
      col.names = c("State", 
                    "Predicted Trump Win (%)", 
                    "Predicted Harris Win (%)", 
                    "Predicted Winner")) %>%
  kable_styling("striped", full_width = F)
```

## US Map Showing Predicted Winner by State on Election Day
To visualize the colour distribution of the predicted results, we emloy the US map [@usmap] to display the results from @tbl-state-probabilities.

@@

## Uncertainty Beyond Data and Modeling
Beyond data and modeling, external factors such as recency bias, social desirability bias, and non-response bias can influence the election outcome. Recency bias significantly shapes public perception, with many expecting Trump to win due to his 2016 victory and narrow 2020 loss, despite lagging in polls. This bias highlights a tendency to focus on recent events while overlooking broader trends. For example, in 2012, Barack Obama outperformed polls, a fact often overshadowed by recent election narratives [@discussion-article3]. This can skew how people respond to polls and how pollsters interpret results, introducing potential polling errors that our model cannot fully account for.

Social desirability bias and non-response bias add further uncertainty, especially regarding Trump’s support base. Social stigma may deter some voters from expressing conservative support openly, leading to under-reported Trump support in polls. Additionally, Trump’s supporters often have lower social trust, making them less likely to participate in polls altogether [@discussion-article3]. If certain groups systematically avoid participation, polling data may under-represent these voters, resulting in an underestimation of Trump’s support in our model.

On the other hand, Harris faces potential challenges such as the Bradley effect and gender-related bias. As a Black female candidate, Harris may encounter undecided voters leaning against her due to underlying biases, similar to what Hillary Clinton faced in 2016 [@discussion-article3]. If undecided voters ultimately vote against Harris, our model may overestimate her support based on polling data alone, affecting prediction accuracy.

@@

## Model Diagnostics {#sec-model-diagnostics}
To assess the validity and reliability of our Bayesian models for predicting voter support for Kamala Harris and Donald Trump, we conducted posterior predictive checks for both models. These checks are essential for evaluating the models' ability to replicate observed polling data based on the estimated parameters. By comparing the distributions of observed support percentages with the predicted values, we can identify discrepancies and assess the overall fit of our models. Posterior predictive checks help verify model assumptions and provide insights into how well the models capture the underlying voter dynamics. A close alignment between observed and predicted values suggests that our models effectively reflect voter support trends in the electoral context. In contrast, significant deviations may reveal areas for model improvement or highlight limitations in our approach. This diagnostic process aims to enhance the credibility of our findings and deepen our understanding of the electoral landscape as the 2024 Presidential Election approaches.

In addition to the Bayesian Beta regression model, other models such as generalized linear models (GLMs) with logistic regression could be employed to predict voter support; however, these may not adequately capture the non-linear trends inherent in polling data. Alternatively, machine learning approaches like random forests or gradient boosting could provide robust predictions by leveraging complex interactions in the data, but they may sacrifice interpretability compared to the Bayesian framework. Ultimately, the choice of model depends on the trade-off between flexibility, interpretability, and the specific characteristics of the data at hand.

@@

### Strengths of the NYT/Siena College Questionnaires

- They have a comprehensive question design. NYT/Sienna College polls ask a wide range of questions that provides a detailed view of the respondent’s preferences and opinions. It helps produce multi-dimensional data that allows for in-depth analysis and segmentation of voter groups. 
- The questionnaires are tailored to likely voters, as identified through the L-2 voter file. This means that the voters are more likely to be up-to-date with relevant policies proposed by different parties, and have formed their own opinions about them, this improves the predictive reliability of the poll. 
- The questionnaires contain detailed demographic questions about the respondents which allow NYT/Siena College to stratify responses by voter subgroups, identifying varied voter bases and diverse issues.

### Weaknesses of the NYT/Siena College Questionnaires

- The length and complexity of the questionnaires might cause the respondents to be less likely to respond or finish the polls. Respondents may rush through questions or drop out before completion. Respondents may also be prone to moderacy response bias where they have the tendency to choose middle options regardless of question content, as well as response order bias where respondents choose answers based on their order. 
- Questionnaires have a dependence on self-reported likelihood to vote. Self-reported likelihood can be unreliable, especially for individuals whose intentions to vote may change close to the actual election date. This can reduce the accuracy of predictions based on likely voter models. 
- There are limited open-ended questions. Many voters may have mixed opinions on specific issues or policies that cannot be explained in a single-choice answer. However, answers to open-ended questions are often difficult to convert to actionable data, so questionnaires tend to rely on closed-ended questions for scalability. This might restrict the ability to capture complex reasonings behind their choices, which could add extra strain on the time and resources used to interpret the answers.

# Idealized Methodology
Building on our discussion of the New York Times/Siena College Poll, we now present an idealized methodology that aims to enhance the accuracy and reliability of our predictions. This approach incorporates best practices and innovative techniques to ensure a comprehensive prediction of voter behavior in the upcoming 2024 U.S. presidential election. We aim to align these methods closely with real-world considerations to ensure it is efficient and represents the views of a broad population well, within a budget of $100,000. To demonstrate our idealized methodology, the following survey is generated, which can be accessed via the the URL provided in [@US-election-survey] under [References -@sec-references].

## Target Population, Sampling Frame, and Sample
